CorReg: model-based pretreatment for regression with correlated variables. Application in steel industry

### 2014-04-01 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@477 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

When one wants to explain a phenomenon based on some covariates, the first statistical method tried frequently is the linear regression. It provides a predictive model with a good interpretability even  for non-statistician and is simple to learn. Therefore, linear regression is used in nearly all the fields where statistics are made, from industry (ballistic models to calibrate the process) to sociology (predicting some numerical properties of a population).
Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators, even whithout bias from modelisation. 
We note the linear regression model:
\begin{equation}
		Y_{|X}=X\beta + \varepsilon
	\end{equation}
	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the  $n\times 1$ response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ the noise of the regression, with $I_n$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$. The $p\times 1$ vector $\beta$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (OLS) by \cite{saporta2006probabilites} 
	\begin{equation}
		\hat{\beta}=(X'X)^{-1}X'Y.
\end{equation}

@@

%bibliographie	
	Because OLS is the minimum-variance unbiased estimator, penalized methods try to reduce the variance introducing some bias to improve the bias-variance trade-off and get better prediction.
Moreover, real datasets imply many irrelevant variables so we have to use variable selection methods.

@@

	The Least Absolute Shrinkage and Selection Operator (LASSO)\cite{tibshirani1996regression} consists in a shrinkage of the regression coefficients based on a $\lambda$ parametric $L_1$ penalty.
		\begin{equation}
		 \hat{\beta}=\operatorname{argmin} \left\lbrace \parallel Y-X\beta\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\beta\parallel_1\leq \lambda
		\end{equation}	
	 The Least Angle Regression\cite{efron2004least} (LAR) Algorithm offers a very efficient way to obtain the whole LASSO path and is very attractive. It requires only the same order of magnitude of computational effort as OLS applied to the full set of covariates. And it  really selects covariates with coefficients set exactly to 0.
	 But LASSO also faces consistency problems \cite{Zhao2006MSC} when confronted  with correlated covariates. Another limitation of the LASSO is that it preserves at most $n$ predictors (troublesome when in high dimension). Some recent variants of the LASSO does exist for the choice of the penalization coefficient like the adaptative LASSO \cite{zou2006adaptive} or the random LASSO \cite{wang2011random}.
	 \\

	Elastic net\cite{zou2005regularization}	is a method developed to be a compromise between Ridge regression and the LASSO. 
%	Given data set $(Y,X)$ and two parameters $(\lambda_1,\lamda_2)$, define artificial data set $(Y^*,X^*)$ by :
%	\begin{equation}
%		X^*_{(n+p)\times p}=(1+\lambda_2)^{-1/2}\left( X  \sqrt{\lambda_2}I \right), \ \ \  Y^*_{(n+p)}=\left( Y 0 \right)
%\end{equation}		
	Elastic net can be written:
	\begin{equation}
		\hat{\beta}=(1+\lambda_2) \operatorname{argmin}\left\lbrace \parallel Y-X\beta \parallel_2^2 \right\rbrace, \textrm{ subject to } (1-\alpha)\parallel\beta\parallel_1+\alpha\parallel\beta\parallel_2^2\leq t \textrm{ for some } t
	\end{equation}
	where $\alpha=\frac{\lambda_2}{(\lambda_1+\lambda_2)}$. 
	%It seems to give good predictions. 
	But it is based on the grouping effect and if the dataset contains two identical variables they will obtain the same coefficient whereas LASSO will choose between one of them and will then obtain same predictions with a more parsimonious model. 
	\\

	Another way of reducing the dimension is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (OSCAR) \cite{bondell2008simultaneous}.
	The CLusterwise Effect REgression\cite{yengo2012variable} (CLERE) describes the $\beta_j$ no longer as fixed effect parameters but as unobserved independant random variables whith grouped $\beta_j$ following a Gaussian Mixture distribution.

The idea is to hope that the model have a small number of groups of covariates and that the mixture will have few enough components to have a number of parameters to estimate significantly lower than $p$. In such a case, it improves interpretability and ability to yeld reliable prediction with a smaller variance on $\hat{\beta}$. But it requires to suppose having many covariates with the same level of effect on the response variable and seems to stay less efficient in prediction than elastic net. Spike and Slab variable selection \cite{ishwaran2005spike} also relies on Gaussian mixture (the spike and the slab) hypothesis for the $\beta_j$ and gives a subset of covariates (not grouped) on which to compute OLS but has no specific protection against correlations issues.
	~\\	~\\
% Principe de la méthode

@@

	Our work is based on the assumption that if we know that correlations are a problem and if we precisely know the correlations, we could use this knowledge to avoid the problem.
	The idea is to suppose explicitly a linear structure between the covariates. 
	It gives a system of linear regression that can be viewed as a recursive Simultaneous Equation Model (SEM)\cite{davidson1993estimation}. Such a system is easy to interpret but estimation don't take advantage of the explicit structure \cite{TIMM} when the structure is straight forward (recursive SEM).
  	Other methods already rely on linear structure and start to take into account correlations but they only consider covariances between the residuals (SUR) \cite{SURzellner} or covariances between the endogenous variables like SPRING (Structured selection of Primordial Relationships IN the General linear model) \cite{chiquetconf}.

	  In this work, we decide to distinguish the response variable from the other endogenous variables (that are on the left of a regression). Thus we don't have a system of regressions but one regression on our response variable and a system of subregressions (without the response variable). And we consider correlations between the explicative covariates of the main regression, not between the residuals.
	  The structure is supposed to be the source of the correlations and allows us to define a marginal regression model based on a reduced set of independent covariates without significant information loss. It can be seen as a pretreatment on the dataset because other regression methods can the be used on this reduced and correlation-free set covariates. So we may obtain two kinds of zeros in our first model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables. This two kinds of zero won't be interpreted in the same way and thus consistency issues don't mean interpretation issues any more. So we dodge the drawbacks of both grouping effect and variable selection.

>>>>

When one wants to explain a phenomenon based on some covariates, the first statistical method tried frequently is the linear regression. It provides a predictive model with a good interpretability even  for non-statistician and is simple to learn. Therefore, linear regression is used in nearly all the fields where statistics are made, from industry (ballistic models to calibrate the process) to sociology (predicting some numerical properties of a population).
Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators, even whithout bias from modelisation. 
We note the linear regression model:
\begin{equation}
		Y_{|X}=X\beta + \varepsilon
	\end{equation}
	where $X$ is the $n\times p$ matrix of the explicative variables, $Y$ the  $n\times 1$ response vector and $\varepsilon \sim \mathcal{N}(0,\sigma_Y^2I_n)$ the noise of the regression, with $I_n$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$. The $p\times 1$ vector $\beta$ is the vector of the coefficients of the regression, estimated with Ordinary Least Squares (\textsc{OLS} \cite{saporta2006probabilites}) by  
	\begin{equation}
		\hat{\beta}=(X'X)^{-1}X'Y.
\end{equation}

@@

%bibliographie	
	Because \textsc{OLS} is the minimum-variance unbiased estimator, penalized methods try to reduce the variance introducing some bias to improve the bias-variance trade-off and get better prediction.
Moreover, real datasets imply many irrelevant variables so we have to use variable selection methods.

@@

	The Least Absolute Shrinkage and Selection Operator (\textsc{LASSO} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a $\lambda$ parametric $L_1$ penalty.
		\begin{equation}
		 \hat{\beta}=\operatorname{argmin} \left\lbrace \parallel Y-X\beta\parallel_2^2 \right\rbrace \textrm{ subject to } \parallel\beta\parallel_1\leq \lambda
		\end{equation}	
	 The Least Angle Regression (\textsc{LAR} \cite{efron2004least}) Algorithm offers a very efficient way to obtain the whole LASSO path and is very attractive. It requires only the same order of magnitude of computational effort as OLS applied to the full set of covariates. And it  really selects covariates with coefficients set exactly to 0.
	 But LASSO also faces consistency problems \cite{Zhao2006MSC} when confronted  with correlated covariates. Another limitation of the \textsc{LASSO} is that it preserves at most $n$ predictors (troublesome when in high dimension). Some recent variants of the \textsc{LASSO} does exist for the choice of the penalization coefficient like the adaptative \textsc{LASSO} \cite{zou2006adaptive} or the random LASSO \cite{wang2011random}.
	 \\

	Elastic net\cite{zou2005regularization}	is a method developed to be a compromise between Ridge regression and the \textsc{LASSO}. 
%	Given data set $(Y,X)$ and two parameters $(\lambda_1,\lamda_2)$, define artificial data set $(Y^*,X^*)$ by :
%	\begin{equation}
%		X^*_{(n+p)\times p}=(1+\lambda_2)^{-1/2}\left( X  \sqrt{\lambda_2}I \right), \ \ \  Y^*_{(n+p)}=\left( Y 0 \right)
%\end{equation}		
	Elastic net can be written:
	\begin{equation}
		\hat{\beta}=(1+\lambda_2) \operatorname{argmin}\left\lbrace \parallel Y-X\beta \parallel_2^2 \right\rbrace, \textrm{ subject to } (1-\alpha)\parallel\beta\parallel_1+\alpha\parallel\beta\parallel_2^2\leq t \textrm{ for some } t
	\end{equation}
	where $\alpha=\frac{\lambda_2}{(\lambda_1+\lambda_2)}$. 
	%It seems to give good predictions. 
	But it is based on the grouping effect and if the dataset contains two identical variables they will obtain the same coefficient whereas LASSO will choose between one of them and will then obtain same predictions with a more parsimonious model. 
	\\

	Another way of reducing the dimension is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (\textsc{OSCAR} \cite{bondell2008simultaneous}).
	The CLusterwise Effect REgression (\textsc{CLERE} \cite{yengo2012variable}) describes the $\beta_j$ no longer as fixed effect parameters but as unobserved independant random variables whith grouped $\beta_j$ following a Gaussian Mixture distribution.

The idea is to hope that the model have a small number of groups of covariates and that the mixture will have few enough components to have a number of parameters to estimate significantly lower than $p$. In such a case, it improves interpretability and ability to yeld reliable prediction with a smaller variance on $\hat{\beta}$. But it requires to suppose having many covariates with the same level of effect on the response variable and seems to stay less efficient in prediction than elastic net. Spike and Slab variable selection \cite{ishwaran2005spike} also relies on Gaussian mixture (the spike and the slab) hypothesis for the $\beta_j$ and gives a subset of covariates (not grouped) on which to compute \textsc{OLS} but has no specific protection against correlations issues.
	~\\	~\\
% Principe de la méthode

@@

	Our work is based on the assumption that if we know explicitly the source of correlations, we could use this knowledge to avoid the problem.
	The structure obtained gives a system of linear regression that can be viewed as a recursive Simultaneous Equation Model (\textsc{SEM})\cite{davidson1993estimation}. Such a system is easy to interpret but estimation don't take advantage of the explicit structure \cite{TIMM} when the structure is straight forward (recursive \textsc{SEM}).
  	Other methods take into account correlations but they only consider covariances between the residuals \textsc{SUR} (Seemingly Unrelated Regression \cite{SURzellner}) or the endogenous variables like \textsc{SPRING} (Structured selection of Primordial Relationships IN the General linear model \cite{chiquetconf}).

	  We make the choice to distinguish the response variable from the other endogenous variables (that are on the left of a regression). Thus we don't have a system of regressions but one regression on our response variable and a system of sub-regressions (without the response variable). Then we consider correlations between the explicative covariates of the main regression, not between the residuals.
	  The structure is supposed to be the source of the correlations and allows us to define a marginal regression model based on a reduced set of independent covariates without significant information loss. It can be seen as a pretreatment on the dataset because other regression methods can the be used on this reduced and correlation-free set covariates. So we may obtain two kinds of zeros in our first model : coerced zeros due to correlations (redundant information) and estimated ones with classical variable selection methods applied on remaining variables. This two kinds of zero won't be interpreted in the same way and thus consistency issues don't mean interpretation issues any more. So we dodge the drawbacks of both grouping effect and variable selection.
